Sketch Recognition and Understanding

### 2009-09-26 draft/Results.tex
after dr. mahmoud meeting
after the changes in examples
after spell check

 %We divided the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\
%We divided the dataset into four sets each set contains four different users. Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.  The algorithms are then tested using the four sets and the results are averaged.
\begin{figure}[]\centering
\fbox{  \parbox{8cm}{% 
		\includegraphics[scale=0.7]{images/symbolSet.eps}	}
		}
	\caption[Hs-DB Symbol Set]{The symboles Set in \cite{HeloiseBeautification} } 
	\label{fig:symbolSet}
\end{figure}

@@

 \begin{table*}
	\centering
	%\small
%	\scalebox{0.7}{
		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
	\scalebox{0.5}{
		\begin{tabular}{|p{2cm}|c|c|c|c|c|c|c|c|c|c|c|c|c||p{2cm}|c|c|c|c|p{2cm}|}\hline 
		Categorie / classified as ==$>$	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout	&Number of samples per class	&TP count	&FP count	&Fn	&Precision	&Recall ( TP rate )\\ \hline 
ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0	&152	&150	&27	&2	&0.8474	&0.9868\\ \hline 
heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0	&152	&130	&20	&22	&0.86666	&0.8552\\ \hline 
trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0	&150	&147	&50	&3	&0.7461	&0.98\\ \hline 
pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0	&152	&139	&9	&13	&0.9391	&0.91447\\ \hline 
arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1	&151	&125	&25	&26	&0.83333	&0.82781\\ \hline 
hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0	&151	&135	&2	&16	&0.9854	&0.89403\\ \hline 
square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0	&151	&138	&1	&13	&0.9928	&0.91390\\ \hline 
triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0	&151	&148	&20	&3	&0.88095	&0.98013\\ \hline 
cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0	&153	&146	&6	&7	&0.96052	&0.95424\\ \hline  
cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0	&152	&147	&2	&5	&0.98657	&0.967105\\ \hline 
parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0	&150	&107	&4	&43	&0.963	&0.71333\\  \hline 
moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5	&153	&125	&21	&28	&0.8561	&0.8169\\ \hline  
callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145	&151	&145	&0	&6	&1	&0.9602\\  \hline  
% Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline 
%ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0\\  \hline 
%heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0\\  \hline 
%trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0\\  \hline 
%pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0\\   \hline 
%arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1\\   \hline 
%hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0\\  \hline 
%square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0\\   \hline 
%triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0\\  \hline 
%cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0\\   \hline 
%cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0\\  \hline 
%parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0\\  \hline 
%moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5\\   \hline 
%callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145\\  \hline 
 		\end{tabular}
			}
 		%\end{tabular}
 	%}
	\caption{Confusion Matrix of HS-DB}
	\label{tab:ConfusionMatrix}
	%\end{minipage}
\end{table*}

>>>>

 %We divided the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\
%We divided the dataset into four sets each set contains four different users. Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.  The algorithms are then tested using the four sets and the results are averaged.
\begin{figure}[]\centering
\fbox{  \parbox{8cm}{% 
		\includegraphics[scale=0.7]{images/symbolSet.eps}	}
		}
	\caption[Hs-DB Symbol Set]{The symbols Set in \cite{HeloiseBeautification} } 
	\label{fig:symbolSet}
\end{figure}

@@

 \begin{table*}
 \caption{Confusion Matrix of HS-DB}
	\centering
	%\small
%	\scalebox{0.7}{
		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
	\scalebox{0.5}{
		\begin{tabular}{|p{2cm}|c|c|c|c|c|c|c|c|c|c|c|c|c||p{2cm}|c|c|c|c|p{2cm}|}\hline 
		Category / classified as ==$>$	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout	&Number of samples per class	&TP count	&FP count	&Fn	&Precision	&Recall ( TP rate )\\ \hline 
ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0	&152	&150	&27	&2	&0.8474	&0.9868\\ \hline 
heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0	&152	&130	&20	&22	&0.86666	&0.8552\\ \hline 
trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0	&150	&147	&50	&3	&0.7461	&0.98\\ \hline 
pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0	&152	&139	&9	&13	&0.9391	&0.91447\\ \hline 
arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1	&151	&125	&25	&26	&0.83333	&0.82781\\ \hline 
hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0	&151	&135	&2	&16	&0.9854	&0.89403\\ \hline 
square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0	&151	&138	&1	&13	&0.9928	&0.91390\\ \hline 
triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0	&151	&148	&20	&3	&0.88095	&0.98013\\ \hline 
cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0	&153	&146	&6	&7	&0.96052	&0.95424\\ \hline  
cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0	&152	&147	&2	&5	&0.98657	&0.967105\\ \hline 
parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0	&150	&107	&4	&43	&0.963	&0.71333\\  \hline 
moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5	&153	&125	&21	&28	&0.8561	&0.8169\\ \hline  
callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145	&151	&145	&0	&6	&1	&0.9602\\  \hline  
% Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline 
%ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0\\  \hline 
%heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0\\  \hline 
%trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0\\  \hline 
%pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0\\   \hline 
%arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1\\   \hline 
%hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0\\  \hline 
%square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0\\   \hline 
%triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0\\  \hline 
%cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0\\   \hline 
%cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0\\  \hline 
%parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0\\  \hline 
%moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5\\   \hline 
%callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145\\  \hline 
 		\end{tabular}
			}
 		%\end{tabular}
 	%}

	\label{tab:ConfusionMatrix}
	%\end{minipage}
\end{table*}

### 2009-09-26 draft/chapterSurvey.tex
after spell check

 Agar et al. \cite{polygonfeedback31} used the timestamp for each point and compute time difference between every two successive point to help him find corners as he argued that it provides more distinct maximum than the speed information. He mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals while the pointing device is moving. There are no points while the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary. Contrary to speed information, where the user draws with variant speeds which tends to make the speed information include a lot of noise in the data.

@@

  Evolutionary algorithms (e.g. Genatic Algoirthms) explore solution using a population which is evaluated using fitness functions. The algorithms initially start with a randomly chosen population which is used to generate new population using various operators (e.g. cross over, mutation... ).  After each iteration, the population is evaluated using a fitness function and only the strongest are used to generate new population. The use of population helps to avoid stochastic search problem and explores the solution space to find better global optimal solutions, but unfortunately they may not always converge to a global best solution\cite{PSOpattern}.

 Particle Swarm Algorithm\cite{PSOFirst}is a population based stochastic search where each solution is simulated as a  particle that flies around in a multidimensional search space. A fitness function is used to evaluate the goodness of each particle in solution space. But unlike evolutionally algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fittest. Each particle saves a history of the best location in the swarm and its own best location.  At the end each iteration, the particles moves to a new location based on the global best location and it own history of best location. This use of pervious particle locations gives PSO advantage over Evolutionary algorithms where no similar information is used.

\subsection{PSO vs. GA}
 PSO can be viewed as an algorithm between stochastic search and Evolutionary algorithms\cite{PSOpattern}. In general the PSO is a continuous optimization problem but experiments showed it performed as well as GA which is in general a discrete and binary optimization problem. In general, both GA and PSO share many points, e.g, both algorithms generate the initial populuation randomaly, both use a fitness function or value to evaluate the goodness of solutions and the next populations. Nevertheless, PSO dose not have genetic operators like crossover or mutations. Partices are updated by internal comunication among themselfs based on the local and global best. In contrast to GA where chromosomes shares information with each others, PSO shares information using memory from previous iteration in form of global best solution. This leads the population to move as one group to towards an optimal solution (i.e find solution quickly). %which significally

@@

This section will discribe each step of the PSO Algorithm. Figure\ref{fig:BlockPSO} illustrates the DPSO algorithm in Algorithm \ref{DPSOalg} steps in details. Initially, the swarm particles are created and each of which is a binary vector crossponding to candidate solution. Additionally, the velocities of each particle is initialized at random by drawing uniformly real values from the range [0.0 to 1.0]. Then the evolutionary process of particle movement is iterated until maximum number of iteration or a fitness error achieved. At each iteration,  the fitness value of each particle is evaluated, and the particles individual local best and global best is updated. The particle position is adjusted as follow; the velocities is computed using Equation \ref{eq:Swarm}. Equation \ref{eq:vmax} is used to scale the velocity to values between 0 and 1 which will serve as a probability that the corresponding bit takes the value of 1. The particles bit vectors are updated by Equation \ref{eq:descrite} according to their velocity which measure the trajectory to the target solution based on local experience ($lbest_{i}$) and global intelligence ($gbest_{i}$). Before the end of the iteration the particle may move to a refined location based on the refine procedure used. 
\begin{algorithm}
\caption{ Discrete Particle Swarm Algorithm }
\label{DPSOalg}
\begin{verbatim}
Initialize 
  1. Generate N particles with M dimensions (P_i...P_N) 
  2. Generate velocities for each particle for M dimension randomly. 
  3. Repeat till Maximum number of iteration OR ErrorThreshold  
         For each particle P_i do the following 
         3.a   Compute fitness  (fitness)
         3.b   Determine Lbest  
               if (lbest_fitness<fitness)
                  3.b.1   lbest=P_i
         3.c      Determine Gbest.
               if (lGest_fitness<fitness)
                  3.c.1   lGest=P_i
         3.d   Compute the Velocity for each dimension using eq.3.2 
         3.e   Bound Velocity using eq.3.4
         3.f   Update particle values using eq. 3.3
         3.g   Refine the location of the particle
         end .
     end 
4. return Gbest
\end{verbatim}
  \end{algorithm}

>>>>

 Agar et al. \cite{polygonfeedback31} used the time-stamp for each point and compute time difference between every two successive point to help him find corners as he argued that it provides more distinct maximum than the speed information. He mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals while the pointing device is moving. There are no points while the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary. Contrary to speed information, where the user draws with variant speeds which tends to make the speed information include a lot of noise in the data.

@@

  Evolutionary algorithms (e.g. Genetic Algorithms) explore solution using a population which is evaluated using fitness functions. The algorithms initially start with a randomly chosen population which is used to generate new population using various operators (e.g. cross over, mutation... ).  After each iteration, the population is evaluated using a fitness function and only the strongest are used to generate new population. The use of population helps to avoid stochastic search problem and explores the solution space to find better global optimal solutions, but unfortunately they may not always converge to a global best solution\cite{PSOpattern}.

 Particle Swarm Algorithm\cite{PSOFirst}is a population based stochastic search where each solution is simulated as a  particle that flies around in a multidimensional search space. A fitness function is used to evaluate the goodness of each particle in solution space. But unlike evolutionarily algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fittest. Each particle saves a history of the best location in the swarm and its own best location.  At the end each iteration, the particles moves to a new location based on the global best location and it own history of best location. This use of previous particle locations gives PSO advantage over Evolutionary algorithms where no similar information is used.

\subsection{PSO vs. GA}
 PSO can be viewed as an algorithm between stochastic search and Evolutionary algorithms\cite{PSOpattern}. In general the PSO is a continuous optimization problem but experiments showed it performed as well as GA which is in general a discrete and binary optimization problem. In general, both GA and PSO share many points, e.g, both algorithms generate the initial population randomly, both use a fitness function or value to evaluate the goodness of solutions and the next populations. Nevertheless, PSO dose not have genetic operators like crossover or mutations. Particles are updated by internal communication among themselves based on the local and global best. In contrast to GA where chromosomes shares information with each others, PSO shares information using memory from previous iteration in form of global best solution. This leads the population to move as one group to towards an optimal solution (i.e find solution quickly). %which significally

@@

This section will describe each step of the PSO Algorithm. Figure\ref{fig:BlockPSO} illustrates the DPSO algorithm in Algorithm \ref{DPSOalg} steps in details. Initially, the swarm particles are created and each of which is a binary vector corresponding to candidate solution. Additionally, the velocities of each particle is initialized at random by drawing uniformly real values from the range [0.0 to 1.0]. Then the evolutionary process of particle movement is iterated until maximum number of iteration or a fitness error achieved. At each iteration,  the fitness value of each particle is evaluated, and the particles individual local best and global best is updated. The particle position is adjusted as follow; the velocities is computed using Equation \ref{eq:Swarm}. Equation \ref{eq:vmax} is used to scale the velocity to values between 0 and 1 which will serve as a probability that the corresponding bit takes the value of 1. The particles bit vectors are updated by Equation \ref{eq:descrite} according to their velocity which measure the trajectory to the target solution based on local experience ($lbest_{i}$) and global intelligence ($gbest_{i}$). Before the end of the iteration the particle may move to a refined location based on the refine procedure used. 
\begin{algorithm}
\caption{ Discrete Particle Swarm Algorithm }
\label{DPSOalg}
\begin{verbatim}
Initialize 
  1. Generate N particles with M dimensions (P_i...P_N) 
  2. Generate velocities for each particle for M dimension randomly. 
  3. Repeat till Maximum number of iteration OR ErrorThreshold  
         For each particle P_i do the following 
         3.a   Compute fitness  (fitness)
         3.b   Determine Lbest  
               if (lbest_fitness<fitness)
                  3.b.1   lbest=P_i
         3.c      Determine Gbest.
               if (lGest_fitness<fitness)
                  3.c.1   lGest=P_i
         3.d   Compute the Velocity for each dimension using eq.3.2 
         3.e   Bound Velocity using eq.3.4
         3.f   Update particle values using eq. 3.3
         3.g   Refine the location of the particle
         end .
     end 
4. return Gbest
\end{verbatim}
  \end{algorithm}
